---
title: "Style Portfolio Analysis data clean"
output: pdf_document
date: "2024-03-17"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("D:/korean discount project/korea discount group pe analysis")
```

```{r}
library(readxl)
library(ggplot2)
library(knitr)
library(tidyr)
library(dplyr)
library(stringr)
library(car)
library(reshape2)
library(patchwork)
library(purrr)
```

```{r}
# Clear the working directory
rm(list = ls())
```


```{r cars}
selected_regions <-c('China','Hongkong','Japan','Korea','Taiwan','US','Vietnam')

countries <- c("Australia", "Austria", "Belgium", "Brazil", "Canada", "Chile", "China", "France", "Germany", "Hongkong", "India", "Indonesia", "Israel", "Italy", "Japan", "Korea", "Malaysia", "Netherland", "New Zealand", "Philippines", "Singapore", "South Africa", "Spain", "Sweden", "Switzerland", "Taiwan", "Thailand", "UK", "US", "Vietnam")

asia_pacific <- c("Australia", "China", "HongKong", "India", "Indonesia", "Japan", "Korea", "Malaysia", "New Zealand", "Philippines", "Singapore", "Taiwan", "Thailand", "Vietnam")

emerging_markets <- c("Brazil", "China","Chile", "India", "Indonesia", "Malaysia", "Philippines", "South Africa", "Thailand", "Vietnam")

developed_markets <- c("Australia", "Austria", "Belgium", "Canada",  "France", "Korea", "Germany", "HongKong", "Israel", "Italy", "Japan", "Netherland", "New Zealand", "Singapore", "South Africa", "Spain", "Sweden", "Switzerland","Taiwan", "Hongkong","UK", "US")
```


```{r}
date_sequence <- seq(as.Date("1997-03-01"), as.Date("2023-03-01"), by = "month")

# Format the dates to only include year and month (YYYY-MM)
formatted_dates <- format(date_sequence, "%Y-%m")
```

Run summary statistics.R script here, which will summarize the sample description

```{r}
table1_data <- read.csv("results/table1.csv")
```

Summary statistics for selected regions from 1997-03 to 2023-03
```{r}
colnames(table1_data)[1] <- "Country"

current_col_names <- colnames(table1_data)[-1]  # Exclude the first column (Country)

# Remove the "X" prefix and convert to a standard format (e.g., "1996.09" to "1996-09")
dates_in_cols <- gsub("X", "", current_col_names)
dates_in_cols <- gsub("\\.", "-", dates_in_cols)
# Convert the dates to Date objects for comparison
dates_as_date <- as.Date(paste0(dates_in_cols, "-01"), format = "%Y-%m-%d")

# Define the range of dates to keep
start_date <- as.Date("1997-03-01")
end_date <- as.Date("2023-03-01")

# Filter columns that fall within the date range
cols_to_keep <- which(dates_as_date >= start_date & dates_as_date <= end_date)

# Keep only the "Country" column and the columns within the date range
table1_data <-table1_data[, c(1, cols_to_keep + 1)]  # +1 because we excluded the first column earlier
colnames(table1_data)[-1] <- formatted_dates

summary_statistics = table1_data %>% filter(Country %in% selected_regions)
summary_statistics 
```

this is created from data cleaning file
```{r}
merged_data<- read.csv("results/merged_data.csv")
```

```{r}
forward_pe_stats <- merged_data %>%
  summarise(
    Q1 = quantile(forward_PE, 0.25, na.rm = TRUE),
    Q3 = quantile(forward_PE, 0.75, na.rm = TRUE),
    IQR = Q3 - Q1,
    Lower_Bound = Q1 - 1.5 * IQR,
    Upper_Bound = Q3 + 1.5 * IQR
  )

# Filter out both lower and higher outliers
merged_data_clean <- merged_data %>%
  filter(
    is.na(forward_PE) |  # Keep NA values if any
    (forward_PE >= forward_pe_stats$Lower_Bound & forward_PE <= forward_pe_stats$Upper_Bound)
  )

# Step 3: Report the results
cat("Original number of observations:", nrow(merged_data), "\n")
cat("Number of observations removed (outliers):", nrow(merged_data) - nrow(merged_data_clean), "\n")
cat("Remaining observations after removing outliers:", nrow(merged_data_clean), "\n")
```

```{r}
# List of metrics to sort on
metrics <- c("Profitability", "AssetGrowth", "MarketCap", "PB", "Momentum_12M")

# Nested lists to store results for all countries
all_portfolio_dicts <- list()
all_date_pe_dfs <- list()

# Loop over each country
for (country in selected_regions) {
  cat("Processing country:", country, "\n")
  
  # Step 1: Filter merged_data for the current country
  country_test <- merged_data_clean %>%
    filter(Country == country)
  
  # Skip if no data for the country
  if (nrow(country_test) == 0) {
    cat("No data found for", country, ". Skipping...\n")
    next
  }
  
  # Step 2: Calculate 12-month momentum
  country_test <- country_test %>%
    group_by(Type) %>%  # Group by firm (Type)
    mutate(
      Price_12M_Ago = lag(price, n = 12, order_by = Date),
      Momentum_12M = if_else(
        !is.na(Price_12M_Ago) & Price_12M_Ago != 0,
        (price / Price_12M_Ago) - 1,
        NA_real_
      )
    ) %>%
    ungroup() %>%
    select(-Price_12M_Ago)
  
  # Step 3: Process each metric for the current country
  portfolio_dicts <- list()
  date_pe_dfs <- list()
  
  for (metric in metrics) {
    cat("  Processing metric:", metric, "\n")
    
    # Ensure the data is clean (remove rows with NA for the current metric)
    country_test_clean <- country_test %>%
      filter(!is.na(.data[[metric]]))
    
    # Skip if no data remains after filtering
    if (nrow(country_test_clean) == 0) {
      cat("  No valid data for", metric, "in", country, ". Skipping...\n")
      next
    }
    
    # Group by Date and create quantiles based on the current metric
    portfolio_data <- country_test_clean %>%
      group_by(Date) %>%
      arrange(.data[[metric]]) %>%  # Sort by the current metric
      mutate(
        Quantile = ntile(.data[[metric]], 3),  # Create tertiles (3 quantiles)
        Quantile_Label = case_when(
          Quantile == 1 ~ "Low",
          Quantile == 2 ~ "Medium",
          Quantile == 3 ~ "High",
          TRUE ~ NA_character_
        )
      ) %>%
      ungroup()
    
    # Create a nested list (dictionary) for portfolio positions
    portfolio_dict <- list()
    unique_dates <- unique(portfolio_data$Date)
    
    for (date in unique_dates) {
      date_data <- portfolio_data %>%
        filter(Date == date)
      
      total_firms <- nrow(date_data)
      
      if (total_firms >= 10) {
        high_firms <- date_data %>%
          filter(Quantile_Label == "High") %>%
          pull(Type)
        
        medium_firms <- date_data %>%
          filter(Quantile_Label == "Medium") %>%
          pull(Type)
        
        low_firms <- date_data %>%
          filter(Quantile_Label == "Low") %>%
          pull(Type)
        
        date_quantiles <- list(
          High = high_firms,
          Medium = medium_firms,
          Low = low_firms
        )
        
        portfolio_dict[[as.character(date)]] <- date_quantiles
      }
    }
    
    # Store the portfolio dictionary for this metric
    portfolio_dicts[[metric]] <- portfolio_dict
    
    # Calculate average forward_PE for each date and quantile
    date_pe_averages <- list()
    
    for (date in names(portfolio_dict)) {
      high_firms <- portfolio_dict[[date]]$High
      medium_firms <- portfolio_dict[[date]]$Medium
      low_firms <- portfolio_dict[[date]]$Low
      
      date_data <- country_test %>%
        filter(Date == date)
      
      high_avg_pe <- if (length(high_firms) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% high_firms], na.rm = TRUE)
      } else {
        NA_real_
      }
      
      medium_avg_pe <- if (length(medium_firms) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% medium_firms], na.rm = TRUE)
      } else {
        NA_real_
      }
      
      low_avg_pe <- if (length(low_firms) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% low_firms], na.rm = TRUE)
      } else {
        NA_real_
      }
      
      date_pe_averages[[date]] <- list(
        High = high_avg_pe,
        Medium = medium_avg_pe,
        Low = low_avg_pe
      )
    }
    
    # Convert to a dataframe
    date_pe_df <- do.call(rbind, lapply(names(date_pe_averages), function(date) {
      data.frame(
        Date = date,
        High_PE = date_pe_averages[[date]]$High,
        Medium_PE = date_pe_averages[[date]]$Medium,
        Low_PE = date_pe_averages[[date]]$Low,
        stringsAsFactors = FALSE
      )
    })) %>%
      arrange(Date)
    
    # Reset row names
    rownames(date_pe_df) <- NULL
    
    # Store the dataframe for this metric
    date_pe_dfs[[metric]] <- date_pe_df
  }
  
  # Store results for the current country
  all_portfolio_dicts[[country]] <- portfolio_dicts
  all_date_pe_dfs[[country]] <- date_pe_dfs
  
  # Print summary for the country
  cat("Completed processing for", country, "\n")
  cat("Metrics processed:", paste(names(date_pe_dfs), collapse = ", "), "\n\n")
}

```
all_date_pe_dfs record the PE values in time dimension of different style portfolio in different countries
all_portfolio_dicts record the dictionary for the portfolio construction




```{r}

# Loop over each country
for (country in selected_regions) {
  cat("Processing country:", country, "\n")
  
  # Extract country-specific data from all_date_pe_dfs
  country_date_pe_dfs <- all_date_pe_dfs[[country]]
  
  # Check if data exists for this country
  if (is.null(country_date_pe_dfs) || length(country_date_pe_dfs) == 0) {
    cat("No data available for", country, "in all_date_pe_dfs. Skipping...\n")
    next
  }
  
  # Ensure Date column is of type Date for each metric's dataframe
  country_date_pe_dfs <- lapply(country_date_pe_dfs, function(df) {
    if (!is.null(df) && "Date" %in% colnames(df)) {
      # Convert Date from "YYYY-MM" to Date type by appending "-01"
      if (is.character(df$Date)) {
        df$Date <- as.Date(paste0(df$Date, "-01"), format = "%Y-%m-%d")
      } else {
        df$Date <- as.Date(df$Date)  # If already a Date-like object, ensure it's Date type
      }
    }
    return(df)
  })
  
  # Loop over each metric and create individual plots
  for (metric in metrics) {
    cat("  Generating plot for metric:", metric, "\n")
    
    # Check if data exists for this metric
    if (is.null(country_date_pe_dfs[[metric]]) || nrow(country_date_pe_dfs[[metric]]) == 0) {
      cat("    No data for", metric, "in", country, ". Skipping...\n")
      next
    }
    
    # Reshape the data to long format
    date_pe_long <- country_date_pe_dfs[[metric]] %>%
      pivot_longer(
        cols = c(High_PE, Medium_PE, Low_PE),
        names_to = "Quantile",
        values_to = "Average_PE"
      ) %>%
      mutate(
        Quantile = case_when(
          Quantile == "High_PE" ~ "High",
          Quantile == "Medium_PE" ~ "Medium",
          Quantile == "Low_PE" ~ "Low",
          TRUE ~ Quantile
        )
      )
    
    # Create the ggplot object
    p <- ggplot(data = date_pe_long, aes(x = Date, y = Average_PE, color = Quantile)) +
      geom_line(size = 1) +
      geom_point(size = 2) +
      labs(
        title = paste("Time Series of Average Forward PE by Quantile (Sorted by", metric, ") -", country),
        x = "Date",
        y = "Average Forward PE",
        color = "Quantile"
      ) +
      theme_minimal() +
      theme(
        plot.title = element_text(hjust = 0.5),
        legend.position = "top"
      )
    
    # Print the plot
    print(p)
    
    # Save the plot
    ggsave(
      filename = paste0("results/countries_pe/forward_pe_time_series_", country, "_", metric, ".png"),
      plot = p,
      width = 10,
      height = 6,
      dpi = 300
    )
  }
  
  cat("Plotting completed for", country, "\n\n")
}

cat("Plotting completed for all countries.\n")
```

```{r}
em_countries <- intersect(selected_regions, emerging_markets)
dm_countries <- intersect(selected_regions, developed_markets)

# Print results
cat("Countries in selected_regions that are Emerging Markets (EM):\n")
print(em_countries)

cat("Countries in selected_regions that are Developed Markets (DM):\n")
print(dm_countries)

```
```{r}
grouped_portfolio_dicts <- list()
# Get all unique dates across all regions and metrics
all_dates <- unique(unlist(lapply(all_portfolio_dicts, function(region_dict) {
  lapply(region_dict, function(metric_dict) {
    names(metric_dict)
  })
})))

# Loop over each metric
for (metric in metrics) {
  cat("Processing metric:", metric, "\n")
  
  # Initialize sublist for this metric
  grouped_portfolio_dicts[[metric]] <- list()
  
  # Loop over each date
  for (date in all_dates) {
    # Initialize portfolios for EM and DM
    em_high <- c()
    em_medium <- c()
    em_low <- c()
    dm_high <- c()
    dm_medium <- c()
    dm_low <- c()
    
    # Aggregate stocks from EM and DM countries
    for (region in selected_regions) {
      if (!is.null(all_portfolio_dicts[[region]][[metric]][[date]])) {
        stocks <- all_portfolio_dicts[[region]][[metric]][[date]]
        
        if (region %in% em_countries) {
          em_high <- c(em_high, stocks$High)
          em_medium <- c(em_medium, stocks$Medium)
          em_low <- c(em_low, stocks$Low)
        } else if (region %in% dm_countries) {
          dm_high <- c(dm_high, stocks$High)
          dm_medium <- c(dm_medium, stocks$Medium)
          dm_low <- c(dm_low, stocks$Low)
        }
      }
    }
    
    # Store the grouped portfolios for this date
    grouped_portfolio_dicts[[metric]][[date]] <- list(
      EM_High = em_high,
      EM_Medium = em_medium,
      EM_Low = em_low,
      DM_High = dm_high,
      DM_Medium = dm_medium,
      DM_Low = dm_low
    )
  }
}
```

grouped_portfolio_dicts record the dictionary for the portfolio construction in developed and emerging market

```{r}
# Step 2: Create a separate dataframe for PE averages
portfolio_pe_df <- do.call(rbind, lapply(metrics, function(metric) {
  cat("Calculating PE averages for metric:", metric, "\n")
  
  do.call(rbind, lapply(all_dates, function(date) {
    # Get the grouped stocks for this metric and date
    portfolios <- grouped_portfolio_dicts[[metric]][[date]]
    
    # Filter merged_data_clean for this date, converting date to match merged_data_clean
    date_data <- merged_data_clean %>%
      filter(Date == date)
    
    # Calculate average PE for each group
    data.frame(
      Metric = metric,
      Date = date,
      EM_High_PE = if (length(portfolios$EM_High) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$EM_High], na.rm = TRUE)
      } else NA_real_,
      EM_Medium_PE = if (length(portfolios$EM_Medium) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$EM_Medium], na.rm = TRUE)
      } else NA_real_,
      EM_Low_PE = if (length(portfolios$EM_Low) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$EM_Low], na.rm = TRUE)
      } else NA_real_,
      DM_High_PE = if (length(portfolios$DM_High) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$DM_High], na.rm = TRUE)
      } else NA_real_,
      DM_Medium_PE = if (length(portfolios$DM_Medium) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$DM_Medium], na.rm = TRUE)
      } else NA_real_,
      DM_Low_PE = if (length(portfolios$DM_Low) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$DM_Low], na.rm = TRUE)
      } else NA_real_,
      stringsAsFactors = FALSE
    )
  }))
})) %>%
  mutate(Date = as.Date(paste0(Date, "-01"), format = "%Y-%m-%d")) %>%
  arrange(Date, Metric)
```
portfolio_pe_df stores the PE averages for DM and EM 


```{r}
# Convert data to long format
portfolio_pe_long <- portfolio_pe_df %>%
  pivot_longer(cols = c("EM_High_PE", "EM_Medium_PE", "EM_Low_PE", "DM_High_PE", "DM_Medium_PE", "DM_Low_PE"),
               names_to = "Category", values_to = "Value") %>%
  mutate(
    Region = ifelse(grepl("EM", Category), "EM", "DM"),
    Level = case_when(
      grepl("High", Category) ~ "High",
      grepl("Medium", Category) ~ "Medium",
      grepl("Low", Category) ~ "Low"
    )
  )

# Define color and linetype mapping
color_mapping <- c("High" = "red", "Medium" = "blue", "Low" = "green")
linetype_mapping <- c("EM" = "dashed", "DM" = "solid")

# Generate separate plots for each metric
unique_metrics <- unique(portfolio_pe_long$Metric)

for (metric in unique_metrics) {
  plot <- ggplot(portfolio_pe_long %>% filter(Metric == metric), 
                 aes(x = Date, y = Value, color = Level, linetype = Region)) +
    geom_line() +
    scale_color_manual(values = color_mapping) +
    scale_linetype_manual(values = linetype_mapping) +
    theme_minimal() +
  
    labs(title = paste(" Time Series of Average Forward PE by Quantile Sorted by", metric), y = "Value", x = "Date", color = "Level", linetype = "Region")
  
  print(plot)  # Display the plot
}
```

The PE for China is quite high, that is why the EM is high.

Now I create two group to exclude the China effect
Group 1 (DM): US and Japan
Group 2 (EM): Vietnam, Hongkong and Korea 
This measure is according to the Korea Discount paper

```{r}
group1<-c('Vietnam','Hongkong','Korea')
group2<-c('Japan','US')

grouped_portfolio_dicts_v2 <- list()
# Get all unique dates across all regions and metrics
all_dates <- unique(unlist(lapply(all_portfolio_dicts, function(region_dict) {
  lapply(region_dict, function(metric_dict) {
    names(metric_dict)
  })
})))

# Loop over each metric
for (metric in metrics) {
  cat("Processing metric:", metric, "\n")
  
  # Initialize sublist for this metric
  grouped_portfolio_dicts[[metric]] <- list()
  
  # Loop over each date
  for (date in all_dates) {
    # Initialize portfolios for EM and DM
    em_high <- c()
    em_medium <- c()
    em_low <- c()
    dm_high <- c()
    dm_medium <- c()
    dm_low <- c()
    
    # Aggregate stocks from EM and DM countries
    for (region in selected_regions) {
      if (!is.null(all_portfolio_dicts[[region]][[metric]][[date]])) {
        stocks <- all_portfolio_dicts[[region]][[metric]][[date]]
        
        if (region %in% group1) {
          em_high <- c(em_high, stocks$High)
          em_medium <- c(em_medium, stocks$Medium)
          em_low <- c(em_low, stocks$Low)
        } else if (region %in% group2) {
          dm_high <- c(dm_high, stocks$High)
          dm_medium <- c(dm_medium, stocks$Medium)
          dm_low <- c(dm_low, stocks$Low)
        }
      }
    }
    
    # Store the grouped portfolios for this date
    grouped_portfolio_dicts_v2[[metric]][[date]] <- list(
      EM_High = em_high,
      EM_Medium = em_medium,
      EM_Low = em_low,
      DM_High = dm_high,
      DM_Medium = dm_medium,
      DM_Low = dm_low
    )
  }
}
```

```{r}
# Step 2: Create a separate dataframe for PE averages
portfolio_pe_df_v2 <- do.call(rbind, lapply(metrics, function(metric) {
  cat("Calculating PE averages for metric:", metric, "\n")
  
  do.call(rbind, lapply(all_dates, function(date) {
    # Get the grouped stocks for this metric and date
    portfolios <- grouped_portfolio_dicts_v2[[metric]][[date]]
    
    # Filter merged_data_clean for this date, converting date to match merged_data_clean
    date_data <- merged_data_clean %>%
      filter(Date == date)
    
    # Calculate average PE for each group
    data.frame(
      Metric = metric,
      Date = date,
      EM_High_PE = if (length(portfolios$EM_High) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$EM_High], na.rm = TRUE)
      } else NA_real_,
      EM_Medium_PE = if (length(portfolios$EM_Medium) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$EM_Medium], na.rm = TRUE)
      } else NA_real_,
      EM_Low_PE = if (length(portfolios$EM_Low) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$EM_Low], na.rm = TRUE)
      } else NA_real_,
      DM_High_PE = if (length(portfolios$DM_High) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$DM_High], na.rm = TRUE)
      } else NA_real_,
      DM_Medium_PE = if (length(portfolios$DM_Medium) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$DM_Medium], na.rm = TRUE)
      } else NA_real_,
      DM_Low_PE = if (length(portfolios$DM_Low) > 0) {
        mean(date_data$forward_PE[date_data$Type %in% portfolios$DM_Low], na.rm = TRUE)
      } else NA_real_,
      stringsAsFactors = FALSE
    )
  }))
})) %>%
  mutate(Date = as.Date(paste0(Date, "-01"), format = "%Y-%m-%d")) %>%
  arrange(Date, Metric)
```

```{r}
# Convert data to long format
portfolio_pe_long_v2 <- portfolio_pe_df_v2 %>%
  pivot_longer(cols = c("EM_High_PE", "EM_Medium_PE", "EM_Low_PE", "DM_High_PE", "DM_Medium_PE", "DM_Low_PE"),
               names_to = "Category", values_to = "Value") %>%
  mutate(
    Region = ifelse(grepl("EM", Category), "EM", "DM"),
    Level = case_when(
      grepl("High", Category) ~ "High",
      grepl("Medium", Category) ~ "Medium",
      grepl("Low", Category) ~ "Low"
    )
  )

# Define color and linetype mapping
color_mapping <- c("High" = "red", "Medium" = "blue", "Low" = "green")
linetype_mapping <- c("EM" = "dashed", "DM" = "solid")

# Generate separate plots for each metric
unique_metrics <- unique(portfolio_pe_long_v2$Metric)

for (metric in unique_metrics) {
  plot <- ggplot(portfolio_pe_long_v2 %>% filter(Metric == metric), 
                 aes(x = Date, y = Value, color = Level, linetype = Region)) +
    geom_line() +
    scale_color_manual(values = color_mapping) +
    scale_linetype_manual(values = linetype_mapping) +
    theme_minimal() +
  
    labs(title = paste(" Time Series of Average Forward PE by Quantile Sorted by", metric), y = "Value", x = "Date", color = "Level", linetype = "Region")
  
  print(plot)  # Display the plot
}
```

